---
title: "Plan for Fiordland BRUV FLD25"
author: "Adam Smith"
date: today
format: 
  html:
    self-contained: true
    toc: true
---

## Setup

```{r}
#| output: false
library(googlesheets4)
library(sf)
library(sfheaders)
library(tidyverse)
library(southernMaps)
library(mapview)
library(terra)
library(tidyterra)
library(rcartocolor)
library(ggnewscale)
library(ggspatial)
library(kableExtra)
library(ggrepel)
theme_set(theme_bw())
```

## Load NZ coastline data

... and convert to `sf` object.

```{r}
nztm <- detailed_nz_islands |> st_as_sf()
```

## Marine reserves data

```{r}
load(file = "data/mr_fiordland.Rdata")
```

## Load strata

```{r}
strata <- st_read("data/FLD25_Strata.kml")
```
```{r}
plot(strata)
```

  
## Load mark

```{r}
mark <- png::readPNG("img/mark_blank.png") |> grid::rasterGrob(interpolate=TRUE)
```


## BRUV data

### Load from Google Sheet

... and convert to `sf` object.

```{r}
#| output: false
fld24 <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1IQfLTiWTQnTGynuF7fUu8htJxKZuIVJw4OK6y9qRfU8",
  sheet="FLD24") |> 
  mutate(Depth = coalesce(`Depth sensor`,`Depth sounder`)) |> 
  st_as_sf(coords = c('Longitude','Latitude'), crs=4326) |> 
  st_join(strata["Name"]) |> 
  filter(Quality %in% c("Good", "OK"))
```

Write full dataset to csv file (for copying stratum column into BRUVmetadata -- actually nah)

```{r}
read_sheet(
  "https://docs.google.com/spreadsheets/d/1IQfLTiWTQnTGynuF7fUu8htJxKZuIVJw4OK6y9qRfU8",
  sheet="FLD24") |> 
  mutate(Depth = coalesce(`Depth sensor`,`Depth sounder`)) |> 
  st_as_sf(coords = c('Longitude','Latitude'), crs=4326) |> 
  st_join(strata["Name"]) |> 
  write_csv(file = "data/fld24_strata.csv")
```


### Export data for printing

```{r}
dstrat <- c("DZ1","DZ2","DR2","DZ3","DZ4" )
```


```{r}
decimal_to_dm <- function(decimal_degrees) {
  degrees <- floor(decimal_degrees)
  minutes <- (decimal_degrees - degrees) * 60
  return(paste0(degrees, "° ", formatC(minutes, format = "f", digits = 3), "'"))
}
```

```{r}
fld24_toprint <- fld24 |>
  filter(Quality != "BAD") |> 
  as_data_frame() |> 
  transmute(
    SiteID,
    SetRep = substring(OpCode, 7, 13),
    Loc = Location,
    Stratum = Name |> 
      fct_relevel(dstrat),
    Depth,
    lat = st_coordinates(fld24)[,2] |> round(5),
    lon = st_coordinates(fld24)[,1] |> round(5),
    lat2 = st_coordinates(fld24)[,2] |> decimal_to_dm(),
    lon2 = st_coordinates(fld24)[,1] |> decimal_to_dm()
    
  ) |> 
  arrange(Stratum)
```


```{r}
kable(fld24_toprint)
```

```{r}
# fld24_toprint |> write_csv(file = "data/fld24_toprint.csv")
```



### First mapviews

```{r}
mapview(fld24)
```

```{r}
map_fiordland24 <- fld24 |> 
  mutate(Depth = -Depth) |> 
  mapview(zcol="Depth", layer.name = "BRUV drops (Depth)") +
  mapview(mr_fiordland |> 
            filter(Name %in% c(
              "Moana Uta (Wet Jacket Arm) Marine Reserve",
              "Taumoana (Five Finger Peninsula) Marine Reserve"
              )), 
          layer.name = "Marine Reserves", 
          col.region = "yellow")

map_fiordland24
```

```{r}
saveRDS( map_fiordland24, "maps/map_fiordland24.rds")
mapshot(map_fiordland24, url = "maps/map_fiordland24.html")
```

```{r}
nrow(fld24)

fld24 |> 
  group_by(Quality) |> 
  summarise(n = n())
```

```{r}
fld24 |> 
  group_by(`Repeat site from 2021`) |> 
  summarise(n = n())
```

```{r}
fld24  |>
  filter(Quality != "BAD") |> 
  summarise(
  min = min(Depth),
  max = max(Depth),
  mean = mean(Depth),
  median = median(Depth)
  )
```

### Calculate areas 

```{r}
strata_clipped <- strata |> 
  st_transform(2193) |> 
  st_difference( nztm |> st_union() ) 

plot(strata_clipped)

strata_areas <- strata_clipped |> 
  mutate( Area = ( strata |> st_area() / 1e6 ) |> units::drop_units()) |> 
  select(Name, Area) |> 
  st_drop_geometry()

```

```{r}
strata_samples <- fld24 |> 
  filter(Quality != "Bad") |> 
  group_by(Name) |> 
  summarise(n = n()) |> 
  st_drop_geometry() |> 
  left_join(strata_areas) |> 
  mutate(
    Area = round(Area, 1),
    Sites_per_area = round(n / Area, 3)
    ) 
```
```{r}
strata_samples
```


```{r}
kable(strata_samples)
```

```{r}
fld24 |> filter(Quality != "Bad") |> ggplot() + aes(x=Depth) + geom_histogram() 
```

```{r}
fld24 |>
  filter(Quality != "Bad") |> 
  ggplot() + aes(x=Depth) + geom_histogram() + facet_wrap(vars(Name))
```

## Depths

```{r}
fld24 |> 
  ggplot() +
  aes(x = `Depth sounder`, y = `Depth sensor`) +
  geom_point() +
  geom_abline(slope=1, intercept = 0)
```

## Mappy map map

```{r}
bathy_tamatea <- readRDS("data/bathy_tamatea.rds")
```

```{r}
lim_tamatea_nzmg <- ext( c(2000000, 2042000, 5464000, 5488000) )
```


### Points and strata 2024

```{r}
ggmap_fld24_strata <- ggplot() +
  
  geom_spatraster(data = bathy_tamatea) +
  
  scale_fill_hypso_c(name = "Depth (m)", palette="colombia_bathy") +
  new_scale_fill() +
  
  geom_sf(
    data = strata |> 
      filter(Name %in% dstrat),
    aes(fill = Name),
    colour = 'black',
    alpha = .5
  ) +
  
  geom_sf(data = nztm, fill='grey40', colour=NA) +
  
  geom_sf(data = fld24, size = 2, color = "black") +
  
  geom_sf_text(
    data = strata |> 
      filter(Name %in% dstrat), 
    aes(label = Name), 
    colour = "white",
    nudge_y = c(-1500,0,0,0,0)) +  # Label strata
  
  scale_fill_brewer(name = "Stratum", palette = "Set2")  +

  # scale_fill_viridis_d(name = "Stratum", option = "C", alpha = 0.7) +
  
  scale_x_continuous(limits = lim_tamatea_nzmg[1:2], 
                     expand = c(0, 0)) +
  scale_y_continuous(limits = lim_tamatea_nzmg[3:4], 
                     expand = c(0, 0)) +
  
  annotation_scale(location = "br") +
  
  annotate("text", label = "Sea Through Science Ltd.") +
  
  xlab("") + ylab("")

ggmap_fld24_strata
```
```{r}
ggmap_fld24_strata + 
  annotation_custom(mark, xmin = 2031500, xmax = 2042000, ymin = 5464000, ymax = 5470000)

ggsave("maps/map_fld24_strata.png", width = 10, height = 6, dpi = 300)
```


### Strata with Breaksea too

#### get bbox

```{r}
strata_nzmg <- strata |> 
  st_transform(strata, crs = 27200)

bbox_strata <- strata_nzmg |> 
  st_bbox()

bbox_strata
```


```{r}
#| fig-height: 8
#| fig-width: 8
ggmap_allstrata <- ggplot() +

  # geom_spatraster(data = bathy_tamatea, alpha = 0.1) +
  # 
  # scale_fill_hypso_c(name = "Depth (m)", palette="colombia_bathy") +
  # new_scale_fill() +
  
  geom_sf(
    data = strata_nzmg |> 
    st_transform(strata, crs = 27200),
    aes(fill = Name),
    colour = 'black',
    alpha = .5
  ) +
  
  geom_sf(data = nztm, fill='grey40', colour=NA) +
  
  geom_sf(data = fld24, size = 2, color = "black", alpha = .5) +
  
  geom_sf_text(
    data = strata,
    aes(label = Name),
    colour = "black",
    fontface = "bold",
    nudge_y = c(0,0,0,500,-1500,0,0,0,0,0,0,0,0)
    ) +  # Label strata
  
  # scale_fill_brewer(name = "Stratum", palette = "Set2")  +

  scale_x_continuous(limits = c(bbox_strata["xmin"], bbox_strata["xmax"]), 
                     expand = c(0, 0)) +
  scale_y_continuous(limits = c(bbox_strata["ymin"], bbox_strata["ymax"]), 
                     expand = c(0, 0)) +
  
  annotation_scale(location = "br") +
  
  annotate("text", label = "Sea Through Science Ltd.") +
  
  xlab("") + ylab("") +
  
  theme(legend.position = "none")

ggmap_allstrata
```


```{r}
#| fig-height: 8
#| fig-width: 8
ggmap_allstrata + 
  annotation_custom(mark, xmin = 2031500, xmax = 2042000, ymin = 5464000, ymax = 5470000)

ggsave("maps/map_allstrata.png", width = 8, height = 8, dpi = 300)
ggsave("maps/map_allstrata.pdf", width = 8, height = 8)
```

### Labelled map of 2024 sites

```{r}
#| fig-height: 13
#| fig-width: 18
ggmap_fld24_labels <- 
  ggmap_fld24_strata +
  geom_label_repel(
    data = fld24, # |> 
      # mutate(
      #   X = st_coordinates(fld24)[,1],
      #   Y = st_coordinates(fld24)[,2]
      # ),
    aes(label = SiteID, geometry = geometry), 
    stat = "sf_coordinates",
    # size = 1,
    max.overlaps = 12,
    alpha = 0.6,
    box.padding = 0.05, 
    point.padding = 0.05,
    min.segment.length = 0)
 
ggmap_fld24_labels

```



```{r}
#| fig-height: 13
#| fig-width: 18
ggmap_fld24_labels + 
  annotation_custom(mark, xmin = 2031500, xmax = 2042000, ymin = 5464000, ymax = 5470000)

ggsave("maps/map_fld24_labels.png", width = 18, height = 13, dpi = 300)
ggsave("maps/map_fld24_labels.pdf", width = 18, height = 13)
```


### Points and reserves

```{r}
ggmap_fld24 <- ggplot() +
  
  geom_spatraster(data = bathy_tamatea) +
  
  geom_sf(
    data = mr_fiordland, 
    fill = "yellow", col = NA, alpha = .3
    ) +
  
  geom_sf(data = nztm, fill='grey40', colour=NA) +
  
  geom_sf(
    data = fld24,
    size = 2
  ) +
  
  scale_fill_hypso_c(name = "Depth (m)", palette="colombia_bathy") +
  
  new_scale_fill() +
  
  scale_x_continuous(limits = lim_tamatea_nzmg[1:2], 
                     expand = c(0, 0)) +
  scale_y_continuous(limits = lim_tamatea_nzmg[3:4], 
                     expand = c(0, 0)) +
  
  annotation_scale(location = "br") +
  
  annotate("text", label = "Sea Through Science Ltd.")

ggmap_fld24
```

```{r}
ggmap_fld24 + 
  annotation_custom(mark, xmin = 2031500, xmax = 2042000, ymin = 5464000, ymax = 5470000)

ggsave("maps/map_fld24.png", width = 10, height = 6, dpi = 300)
```

